LCA in R using tidySEM

### 2022-10-16 example_analysis/Daniel_LCA.Rmd
Update 16 October
Changed title and short title. Adjusted abstract, Added keywords. Made the use of "LCA" and "mixture models" more consistent based on whether we refer to the whole group of LCA techniques, or a mixture subset.  Minor adjustment to wording throughout the paper. Shortened lines throughout. Capitalized titles and made them uniform. Adjusted section-subsection-subsubsection depth (# vs ## vs ###).  Removed irrelevant comments. Wrote best practices in visualization (##Visualization).  Added the tutorial (now set to eval = F for knitting, however code chunks do not show when using

Before we fit a series of LCA models, we set a random seed
using set.seed().
The seed number can be any digit.
This is an important step as 
there is some inherent randomness in the LCA computations,
and having the same seed number ensures that
two separate researcher obtain exactly the same results
when fitting LCA models.

@@

```{r}
set.seed(123) # setting seed 
res <- mx_lca(data = df, classes = 1:4) # fitting LCA 
```
# Class enumeration

In class enumeration, 
we want to compare a sequence of LCA models fitted to the data.
To do so, we use `tidySEM::table_fit()` 
with the results object as the input.
This function creates the model fit table formatted for publication.
However, it contains a lot of information on each of the four models.
It may be more helpful to focus on a subset of model fit indices
and classification diagnostics which we do in the subsequent step.

```{r}
fit_table <- table_fit(res) # model fit table
fit_table[ , c("Name", "LL", "AIC", "BIC", "Entropy", "prob_min", "prob_max", "n_min", "n_max")] # a selection from the model fit table
```
The model fit table is important for deciding on the final class solution.
Our subset of fit indices and classification diagnostics includes:

**Name** indicates the $K$-class solution.
In our case, we chose to fit 1 - 4 class solutions,
each is represented by its row in the table.
**LL** is the -2*log-likelihood of each model.
**AIC** and **BIC** are the Akaike 
and the Bayesian Information Criterion, respectively.
**prob_min** and **prob_max** are 
the lowest and the highest posterior class probability
by most likely class membership.
**n_min** and **n_max** are the lowest and the highest
class proportion based on the posterior class probabilities.

In this paper, we discussed several possible strategies 
to select the final class solution.

Here, we apply our own. 
First, we examine the -2*log likelihood 
which falls successively with each added class.

Then, we look at the BIC value
which is fairly close for the one, two, and three-class solutions,
but high for the four-class solution.

As previously stated, classification diagnostics should not be used for model selection,
but they can be used to disqualify certain solutions because they are uninterpretable.
We see that prob_min for the four-class solution is low,
therefore we disqualify this solution.

Out of the remaining three solutions,
we notice that entropy is the highest for the three-class solution,
and it has a satisfactory prob_min and n_min.
Based on this, we can retain the three class solution in model selection.
Note that entropy for the one-class solution will always equal to one,
as it is 100% true that every case is in that class.
Based on the low entropy of the two-class solution,
we eliminate this model.

Finally, when comparing the one and three-class solutions,
we consult the BIC which tells us that the added complexity
of having three classes still explains the data better
than a one-class solution.
Therefore, we select the three-class solution as our final-class solution.

@@

If we want to know the probability of each response option's endoresement 
for each class, we can use `tidySEM::table_prob`.
These are thresholds for ordinal dependent variables in the probability scale.

```{r}
library(ggplot2)
plot_prob(res[[3]]) # plotting the response patterns for the final class solution
table_prob(res[[3]]) # tabulating the response patterns for the final class solution
```
In the plot, we can see the distributions each of the response probabilities on the indicators
for each of the three classes.
For instance, we see in Class 1 the most common response to u2 is 2,
while in Class 2 and Class 3 this is 0.
We can also see that response 1 is a rare response not forming the majority in 
any class.
Class 2 distinguishes itself because the majority scores the response 0 category of u3 and u4,
while in Class 1 and 2 this is not the case.
Class 3 distinguishes itself because the most common response to u3 and u4 is 3.

@@

# Extracting posterior class probabilities

>>>>

Before we fit a series of LCA models, we set a random seed
using `set.seed()`.
The seed number can be any digit.
This is an important step as 
there is some inherent randomness in the LCA computations,
and having the same seed number ensures that
two separate researcher obtain exactly the same results
when fitting LCA models.

@@

```{r}
set.seed(123) # setting seed 
res <- mx_lca(data = df, classes = 1:4) # fitting LCA 
```
# Class Enumeration

In class enumeration, 
we want to compare a sequence of LCA models fitted to the data.
To aid the process, we create a model fit table using `tidySEM::table_fit()` 
with the results object as the input.
As the output contains a lot of information on each of the four fitted models,
we select a subset of helpful model fit indices and classification diagnostics.

```{r}
fit_table <- table_fit(res) # model fit table
fit_table[ , c("Name", "LL", "AIC", "BIC", "Entropy", "prob_min", "prob_max", "n_min", "n_max")] # our selection
```
Our selection of fit indices and classification diagnostics includes:

- **Name** = the $K$-class solution
- **LL** = the -2*log-likelihood of each model
- **AIC** = the Akaike Information Criterion value
- **BIC** = the Bayesian Information Criterion value
- **prob_min** = the lowest posterior class probability
                 by most likely class membership
- **prob_max** = the highest posterior class probability
                 by most likely class membership
- **n_min**  = the lowest class proportion based on the posterior class probabilities
- **n_max** = the highest class proportion based on the posterior class probabilities

We discussed several possible strategies to select the final class solution. 
Here, we apply our own.

To aid our interpretation of the results, 
we create an elbow plot showing the trends in information criteria across four models.

```{r elbowplot, message=F, warning=F}
library(tidyverse) # for data-wrangling
library(ggplot2) # for plots

elbow_plot <- fit_table[ , c("Name", "AIC", "BIC")] # extract ICs
elbow_plot %<>% pivot_longer(cols = c("AIC", "BIC"), names_to = "IC", values_to = "Value") # to long format

ggplot(elbow_plot, aes(x = Name, y = Value, group = IC))+
  geom_point(aes(color = IC))+
  geom_line(aes(color = IC))+
  labs(title="Elbow Plot of Information Criteria per Class Solution", x = "Class Solution", y = " Value")+
  scale_color_manual(name = "Information Criterion", values = c(AIC = 'blue', BIC = 'red'))+
  theme_minimal()

```
From the elbow plot, we see that AIC has a lower penalty for model complexity.
However, we are more interested in the BIC values, which are similar
for the one, two and three-class solutions, but the four-class solution fits significantly worse.
For this reason, we eliminate the four-class solution from the selection process.

Then we examine the model fit table.
As expected, the -2*log likelihood falls successively with each added class.
As previously stated, classification diagnostics should not be used for model selection,
but they can be used to disqualify certain solutions because they are uninterpretable.
We see that prob_min for the four-class solution is low,
knowing that this solution also has a high BIC, we disqualify this solution.

Out of the remaining three solutions,
we notice that entropy is the highest for the three-class solution,
and it has a satisfactory prob_min and n_min.
Based on this, we retain the three class solution in model selection.
Note that entropy for the one-class solution will always equal to one,
as it is 100% true that every case is in that class.
Based on the low entropy of the two-class solution,
we eliminate this model.

Finally, when comparing the one and three-class solutions,
we inspect the information criteria.
For BIC, the one-class solution fits better, but the difference is marginal.
AIC tells us that the added complexity
of having three classes still explains the data better
than a one-class solution. 
Therefore, we select the three-class solution as our final-class solution.

@@

If we want to know the probability of each response option's endorsement 
for each class, we can use `tidySEM::table_prob`.
These are thresholds for ordinal dependent variables in the probability scale.

```{r}
plot_prob(res[[3]]) # plotting the response patterns for the final class solution
table_prob(res[[3]]) # tabulating the response patterns for the final class solution
```
In the plot, we can see the distributions of the response probabilities on the indicators
for each of the three classes.
For instance, we see that in Class 1 the most common response to u2 is 2,
while in Class 2 and Class 3 this is 0.
We can also see that response 1 is a rare response not forming the majority in 
any class.
Class 2 distinguishes itself because the majority scores the response 0 category of u3 and u4,
while in Class 1 and 2 this is not the case.
Class 3 distinguishes itself because the most common response to u3 and u4 is 3.

@@

# Extracting Posterior Class Probabilities
